---
title: Stefan Eng
---

```{r, include=FALSE, echo=FALSE}
library(knitcitations)

bib <- c(
  'cresst_eng2018' = "10.1109/ICSC.2018.00047"
)

cleanbib()
cite_options(hyperlink = 'to.doc', citation_format = 'text', style = 'html')
# Silently cite
citep(bib["cresst_eng2018"])
```

## Education
  - *Current*: [Master’s Programme in Mathematical Sciences](https://www.chalmers.se/en/departments/math/education/university-of-gothenburg/programmes/mathematical-sciences/Pages/default.aspx) (Statistics) at Gothenburg University/Chalmers.
  -  Bachelor of Science in Mathematics & Computer Science at California State University, Northridge

## Work Experience

### Scientific Computing @ [Novartis](https://www.novartis.com/)
*June 2019 - Current*

  - Internship in the Scientific Computing and Consulting (SCC) group within Biostatistics at Novartis in Basel, Switzerland.
  - Building R Shiny web applications for interactive data visualization and exploration of clinical trial data.
  
### Programmer/Analyst @ [CRESST/UCLA](http://cresst.org) 
*August 2016 - June 2018*

  - Worked on research projects involving natural language processing (NLP)
    -  Presented results at IEEE Conference on Semantic Computing `r citet(bib['cresst_eng2018'])`
  - Created an interactive game using Angular2/Ionic to test and measure reasoning and problem
solving where students attempted to solve tasks by programming an on-screen vehicle.
    - Conducted multiple data collections at high schools across Los Angeles where around one hundred
students were involved in the data collection.
    - Event data was analyzed and exploratory analysis was conducted to summarize student’s problem
solving and reasoning abilities.


### Data Scientist/Software Developer @ [JPL](https://www.jpl.nasa.gov)
*June 2013 - May 2016*

  - Managed data processing pipeline for Cassini and Dawn missions which includes Scala/Apache
Spark ingest code and managing Elasticsearch cluster.
  - Wrote parsers to bring old proprietary data formats into common and modern formats. Including
XML parsing, web scraping, and Scala’s parser combinators.
  - Lead a project to extract data, using Apache Tika, from word documents into MongoDB and
presented with a web interface.
  
## Publications

```{r bibliography, results = 'asis', echo = FALSE, cache = FALSE, warning = FALSE}
## Print bibliography
bibliography(style = 'html')
```
