---
title: Dynamic R Markdown Reports with Shiny
author: Stefan Eng
date: '2019-08-31'
slug: dynamic-r-markdown-reports-with-shiny
categories:
  - R
  - development
tags:
  - shiny
  - rmarkdown
  - data-science
output:
  blogdown::html_page:
    toc: no
    fig_width: 5
    fig_height: 5
csl: ../../static/bibtex/acm-sig-proceedings.csl
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## Load frequently used packages for blog posts
library(sessioninfo) # for session_info()
library(here)

## For R images to display well in the RSS feed (disable for local preview)
# knitr::opts_knit$set(base.url = 'stefanengineering.com/post/')

# Don't show code by default
knitr::opts_chunk$set(warning = FALSE, comment = NA, message = FALSE)
knitr::opts_knit$set(global.par = TRUE)

# CB Palette like ESL
cbPalette <- c(
  "#999999", "#E69F00", "#56B4E9", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

TL;DR here is an example application. See the explaination below.

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  downloadButton('download')
)

server <- function(input, output) {
  output$download <- downloadHandler(
    filename = "listing.pdf",
    content = function(f) {
      # Create a new empty environment
      # This allows us to pass in only the relevant variables into the report
      e <- new.env()
      # Pass two data sets into our template
      e$datasets <- list(mtcars, iris)
      # Render the document
      rmarkdown::render('template.Rmd',
                        output_format = rmarkdown::pdf_document(),
                        output_file=f,
                        envir = e)
    }
  )
}

shinyApp(ui = ui, server = server)
```

A common issue I have run into is dynamically generating reports.
In particular, during my summer at Novartis there was lots of demands for report listings that could be generated on the fly from a Shiny application.
Using the great [rmarkdown](https://github.com/rstudio/rmarkdown) package makes it really easy to do this.
My approach is to create a template R markdown file and dynamically generate it using [`rmarkdown::render`](https://www.rdocumentation.org/packages/rmarkdown/versions/1.15/topics/render) in a download handler.

The full application is available here: https://github.com/stefaneng/Shiny-Dynamic-Report-Generation

We first need a R markdown template that we can use for the report generation.
We want to print out some data set using the `pander` package to make nicer formatted tables.

<script src="http://gist-it.appspot.com/http://github.com/stefaneng/Shiny-Dynamic-Report-Generation/blob/master/template.Rmd"></script>

To generate the report from Shiny, we use a [downloadHandler](https://shiny.rstudio.com/reference/shiny/0.14/downloadHandler.html) and render the template using [`rmarkdown::render`](https://www.rdocumentation.org/packages/rmarkdown/versions/1.15/topics/render).
This is called each time the user clicks the download button. In the [full example on github](https://github.com/stefaneng/Shiny-Dynamic-Report-Generation), the user can select a data set to include in the report which is more realistic.
The example given is the simplest to get you going on generating your own dynamic reports.

## Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
session_info()
```
