---
title: Long distance relationship calls analysis
author: Stefan Eng
date: '2019-07-14'
slug: long-distance-relationship-calls-analysis
categories:
  - R
  - personal
tags:
  - R
  - analysis
output:
  blogdown::html_page:
    toc: no
    fig_width: 8.32
    fig_height: 5.64
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
library(scales)
library(sessioninfo)

## For R images to display well in the RSS feed (disable for local preview)
# knitr::opts_knit$set(base.url = 'stefanengineering.com/post/')

# Don't show code by default
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, comment = NA, message = FALSE)
knitr::opts_knit$set(global.par = TRUE)

# CB Palette like ESL
cbPalette <- c(
  "#999999", "#E69F00", "#56B4E9", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# My favorite theme

theme_set(ggplot2::theme_minimal(base_size = 15, base_family = "Arial"))

move_date <- date("2018-08-14")
voice_port <- date("2016-05-24")
first_date <- date("2016-02-14")

andrea_df <- readRDS('~/statistics/hangouts_analysis/data/processed_data/andrea_voice_hangouts.rds')

andrea_dist_df <- andrea_df %>%
  # Only take calls after move date and conversations longer than a minute
  filter(pst_starttime >= move_date & duration_seconds > 60)

# Get the time spent each month. Much more obvious that the time increased.
andrea_dist_df$yearmonth <- format(andrea_dist_df$pst_starttime, "%Y-%m")
month_order <- format(seq(date("2018-08-5"), date("2019-06-5"), by = "month"), "%b %y")

andrea_dist_df$nicemonth <- factor(format(andrea_dist_df$pst_starttime, "%b %y"),
                                   levels = month_order)
```

## Introduction
I have been together with my girlfriend for about three and a half years now.
A few months after we started dating, I switched my telephone number over to Google Voice (if anyone else is interested in this, see the [link here](https://support.google.com/voice/answer/1065667#xferin)).
A side effect of this is that I can access my text and call data from practically our entire relationship.
In September 2018, I started a Master's program in statistics at Gothenburg University, and have been continuing our relationship long-distance.

```{r overview-plot, fig.width=12, fig.height=4, fig.cap="Entire relationship of conversations on Google Voice and Hangouts"}
ggplot(andrea_df) +
  geom_histogram(aes(x = date(pst_starttime), weight = duration_seconds / 3600,fill = source),
                 position = "stack",
                 bins = 50) +
  geom_vline(xintercept = move_date, color = cbPalette[1]) +
  geom_vline(xintercept = first_date, color =  cbPalette[1]) +
  geom_vline(xintercept = voice_port, color = cbPalette[1]) +
  # First date
  annotate("text", x = date("2016-09-01"), y = 21, label = "First date") +
  geom_curve(x = date("2016-07-15"), xend = first_date, y = 21, yend = 10,
             size = 0.4,
             arrow = arrow(length = unit(0.07, "inch")),
             curvature = 0.3) +
  annotate("text", x = date("2017-01-01"), y = 10, label = "Ported number to Voice") +
  geom_curve(x = date("2017-01-01"), xend = voice_port, y = 9, yend = 8,
             size = 0.4,
             arrow = arrow(length = unit(0.07, "inch")),
             curvature = -0.1) +
  annotate("text", x = date("2018-01-01"), y = 21, label = "Started long-distance") +
  geom_curve(x = date("2018-04-10"), xend = move_date, y = 20.5, yend = 15,
             size = 0.4,
             arrow = arrow(length = unit(0.07, "inch")),
             curvature = -0.2) +
  # Annotate with timezone
  geom_text(label="Pacific Time", x = first_date - months(3), y = -4, color = cbPalette[1]) +
  coord_cartesian(clip = 'off') +
  labs(x = NULL, y = "Hours", title = "Our relationship according to Google") +
  theme(text = element_text(size = 18),
        panel.grid.major  = element_blank(),
        panel.grid.minor  = element_blank(),
        plot.caption = element_text(size = 9, color = "gray50"),
        legend.position = c(-0.07,0.2),
        legend.title = element_blank(),
        legend.text = element_text(size = 11),
        axis.title.y = element_text(
          angle = 0, hjust = 1, vjust = 0.8, margin = margin(l = 24, r = 10)
        )
  ) +
  scale_fill_manual(values = cbPalette[-1]) +
  scale_x_date(labels = date_format("%b %y"), breaks = "6 month")
```

In \@ref(fig:overview-plot) we show a stacked histogram colored by the service we communicated with (either Google Voice or Hangouts).
This is a combination of both video and voice calls.

```{r per-month-hours, fig.cap="Hours spent talking per month"}
andrea_dist_df %>%
  group_by(nicemonth) %>%
  summarise(n = n(), hours = sum(duration_seconds, na.rm = TRUE) / 3600) %>%
  mutate(visiting = case_when(nicemonth == "Nov 18" ~ .7,
                              nicemonth == "Apr 19" ~ .7,
                              TRUE ~ 1)) %>%
  ggplot(.) +
  geom_bar(aes(x = nicemonth, y = hours, alpha = visiting), stat = "identity", fill = cbPalette[3]) +
  # Andrea visits for 2 weeks in november
  annotate("text", x = 4, y = 25, label = "2 week visit") +
  geom_curve(x = 4, xend = 4, y = 23, yend = 18,
             size = 0.4,
             arrow = arrow(length = unit(0.07, "inch")),
             curvature = -0.1) +
  # I come home for two weeks
  annotate("text", x = 10, y = 28, label = "2 week visit") +
  geom_curve(x = 9, xend = 9, y = 28, yend = 18,
             size = 0.4,
             arrow = arrow(length = unit(0.07, "inch"))) +
  labs(x = "", y = "Hours", title = "Long-distance hours per month") +
  theme(text = element_text(size = 18),
        panel.grid.major.x  = element_blank(),
        panel.grid.minor  = element_blank(),
        axis.title.y = element_text(
          angle = 0, hjust = 1, vjust = 0.8,
          size = 14
        ),
        legend.text = element_text(
          size = 10
        ),
        legend.position = c(.86,.95),
        legend.key.height = unit(0, 'cm'),
        legend.background = element_rect(
          color = "grey90",
          size = .5

        )
  ) +
  guides(alpha = FALSE) +
  scale_alpha(range = c(.5, 1.0))
```

In Figure \@ref(fig:per-month-hours) we see a big increase in number of hours spent from January to March.
This was my first winter and Sweden and the dark and grey was pretty tough.
It was tough being apart during these times so we spent more time on video calls.

```{r call-duration-time, fig.cap="Histogram of hours spent on each conversation"}
# Max conversation time
max_td <- seconds_to_period(max(andrea_dist_df$duration_seconds))
# Nicely formated
max_td_str <- sprintf("%2d hours and %2d minutes", hour(max_td), minute(max_td))

# Histogram of durations
ggplot(andrea_dist_df) +
  geom_histogram(aes(x = duration_seconds / 3600), bins = 30, fill = cbPalette[3]) +
  annotate("text", x = 5, y = 10, label = paste0("Longest call: \n", max_td_str)) +
  geom_curve(x = 5.25, xend =5.95, y = 6, yend = .5,
             size = 0.4,
             arrow = arrow(length = unit(0.07, "inch")),
             curvature = 0.3) +
  labs(x = "Hours", y = "Count", title = "Hours spent per phone call") +
  theme(text = element_text(size = 18),
        panel.grid.minor  = element_blank(),
        axis.title.y = element_text(
          angle = 0, hjust = 1, vjust = 0.9,
          size = 14
        )
  ) +
  scale_x_continuous(breaks = seq(0, 6, length.out = 7))
```

In Figure \@ref(fig:call-duration-time) a histogram of the hours spent on each call are shown.
The distribution is right-skewed with quite a few really long conversations.
Note that this does not take into account conversations that were split into multiple calls.

```{r weekday-hours, fig.cap="Hours spent per week"}
andrea_dist_df$weekday <- wday(andrea_dist_df$pst_starttime, label = TRUE, abbr = FALSE, week_start = 1)
andrea_dist_df %>%
  group_by(weekday) %>%
  summarise(duration_hours = sum(duration_seconds) / 3600) %>%
ggplot(.) +
  geom_bar(aes(x = weekday, y = duration_hours), stat = "identity", fill = cbPalette[3]) +
  labs(x = "", y = "Hours", title = "Total hours on each day of the week") +
  theme(text = element_text(size = 18),
        panel.grid.major.x  = element_blank(),
        panel.grid.minor  = element_blank(),
        axis.title.y = element_text(
          angle = 0, hjust = 1, vjust = 0.8,
          size = 14
        )
  )
```

I was curious if there was a big difference in the hours spent on the phone by week day. In Figure \@ref(fig:weekday-hours) we can see that the weekdays are less than the weekends which is fairly expected. Wednesdays and Thursdays were a bit of a mystery as to why they were less than the other days. My girlfriend saw the graph and reminded me that I had an 8am course on Thursdays and Fridays a few quarters (Wednesday and Thursday night for her).

## Closing Thoughts

Overall I am really happy with the way that our relationship has progressed.
Before leaving I never thought I would describe it as progressing when we were not together.
It has been a great learning experience and I have learned to express myself better in our conversations.
Of course it is not easy being away from the one you love.
It takes a lot of work and hours as you can see from the visualizations.
The effort makes a huge difference though and it is possible to keep up and even improve a relationship while long distance.


## Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
session_info()
```
