---
title: Delete rows from Shiny DT::datatable
author: Stefan Eng
date: '2019-07-06'
slug: delete-rows-from-shiny-dt-datatable
categories:
  - development
  - R
tags:
  - R
  - shiny
  - DT::datatable
output:
  blogdown::html_page:
    toc: no
    fig_width: 5
    fig_height: 5
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## Load frequently used packages for blog posts
library('knitcitations') # for citations
library(shiny)

## For R images to display well in the RSS feed (disable for local preview)
# knitr::opts_knit$set(base.url = 'stefanengineering.com/post/')

# Don't show code by default
knitr::opts_chunk$set(warning = FALSE, comment = NA, message = FALSE, eval = F)
knitr::opts_knit$set(global.par = TRUE)

## Load knitcitations with a clean bibliography
cleanbib()
cite_options(hyperlink = 'to.doc', citation_format = 'text', style = 'html',
             cite.style = 'numeric')

bib <- list(
    'blogdown' = citation('blogdown')[2],
    'knitcitations' = citation('knitcitations'),
    'setInputValue' = 'https://shiny.rstudio.com/articles/communicating-with-js.html'
)
```

<!-- Strange bug where I need some font-awesome icon on the page or else they don't show up in the table -->
`r icon('delete')`

[Demo](https://stefaneng.shinyapps.io/DeleteRowsDT/) and [Code](https://github.com/stefaneng/Shiny-DeleteRowsDT)

This is a simple shiny app that adds a delete button for each row in a data table.
Once a row is deleted, it can be put back into the table with the undo button.

The main idea is that we add a new column to the given data frame that consists of an character representation of an [`actionButton`](https://shiny.rstudio.com/reference/shiny/latest/actionButton.html).
Each button has an id that ends with the row number.
We take advantage of the onclick callback and use `Shiny.setInputValue` `r citep(bib[['setInputValue']])` via javascript bind our own input.
In this simple case, we are only sending back the id of the button that we created.
By default, `setInputValue` will not trigger a new event if the value has not change.
This could happen if we delete a row, then undo the deletion and then try to delete the same row.
To make sure that each click triggers the event, we add `{priority: "event"}` as an argument.
```{js}
Shiny.setInputValue("deletePressed", this.id, {priority: "event"})
```

Now to add the buttons to the column, we use the following function

```{r, eval = T}
#' A column of delete buttons for each row in the data frame for the first column
#'
#' @param df data frame
#' @param id id prefix to add to each actionButton. The buttons will be id'd as id_INDEX.
#' @return A DT::datatable with escaping turned off that has the delete buttons in the first column and \code{df} in the other
deleteButtonColumn <- function(df, id, ...) {
  # function to create one action button as string
  f <- function(i) {
    as.character(
      actionButton(
        # The id prefix with index
        paste(id, i, sep="_"),
        label = NULL,
        icon = icon('trash'),
        onclick = 'Shiny.setInputValue(\"deletePressed\", this.id, {priority: "event"})'))
  }

  deleteCol <- unlist(lapply(seq_len(nrow(df)), f))

  # Return a data table
  DT::datatable(cbind(delete = deleteCol, df),
                # Need to disable escaping for html as string to work
                escape = FALSE,
                options = list(
                  # Disable sorting for the delete column
                  columnDefs = list(
                    list(targets = 1, sortable = FALSE))
                ))
}
```

We can use it as follows:
```{r, eval =T}
df <- data.frame(matrix(sample(1:20, 10), ncol = 2))
deleteButtonColumn(df, 'button')
```

Now in the server code, we have an event `input$deletePressed` that we can react to.
The value is the id of the button that we clicked on, which will be `button_1` if we clicked on button 1.
```{r}
server <- function(input, output) {
  # ...  
  
  observeEvent(input$deletePressed, {
    # Handle delete pressed
  })
}
```

We then need to parse the index from `input$deletePressed`
```{r}
#' Extracts the row id number from the id string
#' @param idstr the id string formated as id_INDEX
#' @return INDEX from the id string id_INDEX
parseDeleteEvent <- function(idstr) {
  res <- as.integer(sub(".*_([0-9]+)", "\\1", idstr))
  if (! is.na(res)) res
}
```

Then we can delete the row by
```{r}
server <- function(input, output) {
  rv <- reactiveValues(data = NULL)
      
  observeEvent(input$deletePressed, {
        rowNum <- parseDeleteEvent(input$deletePressed)
        
        # Delete the row from the data frame
        rv$data <- rv$data[-rowNum,]
      })
  
  output$dtable <- DT::renderDataTable(
      # Add the delete button column
      deleteButtonColumn(rv$data, 'delete_button')
    )
}  
```

That's basically it! In the demo application, I implemented an undo function as well.
Check it out in the code below.

## Final Code

<script src="https://gist-it.appspot.com/https://github.com/stefaneng/Shiny-DeleteRowsDT/blob/master/app.R"></script>

### References

```{r bibliography, results = 'asis', echo = FALSE, cache = FALSE, warning = FALSE, eval = T}
## Print bibliography
bibliography(style = 'html')
```

### Reproducibility

### Session info

```
 setting  value                       
 version  R version 3.6.0 (2019-04-26)
 os       macOS Mojave 10.14.3        
 system   x86_64, darwin15.6.0        
 ui       RStudio                     
 language (EN)                        
 collate  en_US.UTF-8                 
 ctype    en_US.UTF-8                 
 tz       Europe/Stockholm            
 date     2019-07-06     
```

### Packages
```
package      * version  date       lib source        
 assertthat     0.2.1    2019-03-21 [1] CRAN (R 3.6.0)
 backports      1.1.4    2019-04-10 [1] CRAN (R 3.6.0)
 BH             1.69.0-1 2019-01-07 [1] CRAN (R 3.6.0)
 cli            1.1.0    2019-03-19 [1] CRAN (R 3.6.0)
 colorspace     1.4-1    2019-03-18 [1] CRAN (R 3.6.0)
 crayon         1.3.4    2017-09-16 [1] CRAN (R 3.6.0)
 crosstalk      1.0.0    2016-12-21 [1] CRAN (R 3.6.0)
 digest         0.6.20   2019-07-04 [1] CRAN (R 3.6.0)
 DT             0.7      2019-06-11 [1] CRAN (R 3.6.0)
 ellipsis       0.2.0.1  2019-07-02 [1] CRAN (R 3.6.0)
 fansi          0.4.0    2018-10-05 [1] CRAN (R 3.6.0)
 ggplot2        3.2.0    2019-06-16 [1] CRAN (R 3.6.0)
 glue           1.3.1    2019-03-12 [1] CRAN (R 3.6.0)
 gtable         0.3.0    2019-03-25 [1] CRAN (R 3.6.0)
 htmltools      0.3.6    2017-04-28 [1] CRAN (R 3.6.0)
 htmlwidgets    1.3      2018-09-30 [1] CRAN (R 3.6.0)
 httpuv         1.5.1    2019-04-05 [1] CRAN (R 3.6.0)
 jsonlite       1.6      2018-12-07 [1] CRAN (R 3.6.0)
 labeling       0.3      2014-08-23 [1] CRAN (R 3.6.0)
 later          0.8.0    2019-02-11 [1] CRAN (R 3.6.0)
 lattice        0.20-38  2018-11-04 [3] CRAN (R 3.6.0)
 lazyeval       0.2.2    2019-03-15 [1] CRAN (R 3.6.0)
 magrittr       1.5      2014-11-22 [1] CRAN (R 3.6.0)
 MASS           7.3-51.4 2019-03-31 [3] CRAN (R 3.6.0)
 Matrix         1.2-17   2019-03-22 [3] CRAN (R 3.6.0)
 mgcv           1.8-28   2019-03-21 [3] CRAN (R 3.6.0)
 mime           0.7      2019-06-11 [1] CRAN (R 3.6.0)
 munsell        0.5.0    2018-06-12 [1] CRAN (R 3.6.0)
 nlme           3.1-139  2019-04-09 [3] CRAN (R 3.6.0)
 pillar         1.4.2    2019-06-29 [1] CRAN (R 3.6.0)
 pkgconfig      2.0.2    2018-08-16 [1] CRAN (R 3.6.0)
 plyr           1.8.4    2016-06-08 [1] CRAN (R 3.6.0)
 promises       1.0.1    2018-04-13 [1] CRAN (R 3.6.0)
 R6             2.4.0    2019-02-14 [1] CRAN (R 3.6.0)
 RColorBrewer   1.1-2    2014-12-07 [1] CRAN (R 3.6.0)
 Rcpp           1.0.1    2019-03-17 [1] CRAN (R 3.6.0)
 reshape2       1.4.3    2017-12-11 [1] CRAN (R 3.6.0)
 rlang          0.4.0    2019-06-25 [1] CRAN (R 3.6.0)
 scales         1.0.0    2018-08-09 [1] CRAN (R 3.6.0)
 shiny          1.3.2    2019-04-22 [1] CRAN (R 3.6.0)
 sourcetools    0.1.7    2018-04-25 [1] CRAN (R 3.6.0)
 stringi        1.4.3    2019-03-12 [1] CRAN (R 3.6.0)
 stringr        1.4.0    2019-02-10 [1] CRAN (R 3.6.0)
 tibble         2.1.3    2019-06-06 [1] CRAN (R 3.6.0)
 utf8           1.1.4    2018-05-24 [1] CRAN (R 3.6.0)
 vctrs          0.2.0    2019-07-05 [1] CRAN (R 3.6.0)
 viridisLite    0.3.0    2018-02-01 [1] CRAN (R 3.6.0)
 withr          2.1.2    2018-03-15 [1] CRAN (R 3.6.0)
 xtable         1.8-4    2019-04-21 [1] CRAN (R 3.6.0)
 yaml           2.2.0    2018-07-25 [1] CRAN (R 3.6.0)
 zeallot        0.1.0    2018-01-28 [1] CRAN (R 3.6.0)
```
