---
draft: true
title: Long distance relationship calls analysis
author: Stefan Eng
date: '2019-07-14'
slug: long-distance-relationship-calls-analysis
categories:
  - R
  - personal
tags:
  - R
  - analysis
output:
  blogdown::html_page:
    toc: no
    fig_width: 5
    fig_height: 5
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
library(lubridate)
library(sessioninfo)

## For R images to display well in the RSS feed (disable for local preview)
# knitr::opts_knit$set(base.url = 'stefanengineering.com/post/')

# Don't show code by default
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, comment = NA, message = FALSE)
knitr::opts_knit$set(global.par = TRUE)

# CB Palette like ESL
cbPalette <- c(
  "#999999", "#E69F00", "#56B4E9", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# My favorite theme

theme_set(ggplot2::theme_minimal(base_size = 15, base_family = "Arial"))

move_date <- date("2018-08-14")
voice_port <- date("2016-05-24")
first_date <- date("2016-02-14")

andrea_df <- readRDS('~/statistics/hangouts_analysis/data/processed_data/andrea_voice_hangouts.rds')
```

## Introduction
I have been together with my girlfriend for about three and a half years now.
A few months after we started dating, I switched my telephone number over to Google Voice (if anyone else is interested in this, see the [link here](https://support.google.com/voice/answer/1065667#xferin)).
A side effect of this is that I can access my text and call data from practically our entire relationship.
In September 2018, I started a Master's program in statistics at Gothenburg University, and have been continuing our relationship long-distance.

## Google Voice/Hangouts
Overall it has been a good experience and I can keep my same number in whatever country I am in.
As with other Google services, you can download your own data from [takeout.google.com](https://takeout.google.com/settings/takeout).


### Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
session_info()
```
