---
title: test
author: Stefan Eng
date: '2019-06-26'
slug: test
categories: []
tags: []
output:
  blogdown::html_page:
    toc: no
    fig_width: 5
    fig_height: 5
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## For R images to display well in the RSS feed (disable for local preview)
# knitr::opts_knit$set(base.url = 'stefanengineering.com/post/')

# Don't show code by default
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, comment = NA, message = FALSE)
knitr::opts_knit$set(global.par = TRUE)

# CB Palette like ESL
cbPalette <- c(
  "#999999", "#E69F00", "#56B4E9", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# My favorite theme
ggplot2::theme_set(ggplot2::theme_minimal())

## Load frequently used packages for blog posts
library('knitcitations') # for citations
library('sessioninfo') # for session_info()

## Load knitcitations with a clean bibliography
cleanbib()
cite_options(hyperlink = 'to.doc', citation_format = 'text', style = 'html')

bib <- c(
    'blogdown' = citation('blogdown')[2],
    'knitcitations' = citation('knitcitations'),
    'sessioninfo' = citation('sessioninfo')
)
```


### Post content

Typical location to start editing since the bibliography chunk is hidden. Make sure that you selected `R Markdown (.Rmd)` as the _format_ option of the post when using the `New Post` `blogdown` addin.

### Acknowledgements


This blog post was made possible thanks to:

* `r citep(bib[['blogdown']])`
* `r citep(bib[['knitcitations']])`
* `r citep(bib[['sessioninfo']])`

### References

```{r bibliography, results = 'asis', echo = FALSE, cache = FALSE, warning = FALSE}
## Print bibliography
bibliography(style = 'html')
```

### Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
session_info()
```
