---
draft: true
title: Tower property and Martingales
author: Stefan Eng
date: '2020-05-21'
slug: tower-property-and-martingales
categories:
  - math
tags:
  - probability
output:
  blogdown::html_page:
    toc: no
    fig_width: 5
    fig_height: 5
link-citations: yes
csl: ../../static/bibtex/acm-sig-proceedings.csl
bibliography: ../../static/bibtex/probability.bib
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## Load frequently used packages for blog posts
library(sessioninfo) # for session_info()
library(ggplot2)
library(here)

## For R images to display well in the RSS feed (disable for local preview)
# knitr::opts_knit$set(base.url = 'stefanengineering.com/post/')

# Don't show code by default
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, comment = NA, message = FALSE)
knitr::opts_knit$set(global.par = TRUE)

# CB Palette like ESL
cbPalette <- c(
  "#999999", "#E69F00", "#56B4E9", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# My favorite theme
theme_set(ggplot2::theme_minimal(base_size = 15, base_family = "Arial"))
```

## Martingales

A sequence $\{Y_n : n \geq 0\}$ is a **martingale** with respect to $\{X_n : n \geq 0\}$ if for all $n \geq 0$,
  1. $E[|Y_n|] < \infty$
  2. $E[Y_{n + 1} | X_1, \ldots, X_n] = Y_n$
  
If $E[Y_{n + 1} | X_1, \ldots, X_n] \leq Y_n$ then $\{Y_n\}$ is a **supermartingale** and if $E[Y_{n + 1} | X_1, \ldots, X_n] \geq Y_n$ then $\{Y_n\}$ is a **submartingale** [@grimmett2001].


More generally, first we define a **filtration** as a sequence of sub-sigma-fields $\mathcal F_* = \{\mathcal F_0, \mathcal F_1, \ldots\}$ in which $\mathcal F_n \subseteq \mathcal F_{n + 1}$.
A stochastic process $\{Y_n : n \geq 0\}$ is **adapted** to the filtration $\mathcal F_*$ if $Y_n$ is $\mathcal F_n$-measurable for all $n$.

Let $\mathcal F_*$ be a _filtration_ of the probability space $(\Omega, \mathcal F, P)$.
Assume that $\{Y_n : n \geq 0\}$ is _adapted_ to $\mathcal F_*$.

### Tower Property
$$
E[X] = E[E[X|Y]]
$$
If we condition on $Y_1$, then by conditional probability properties we have
$$
E[E(X | Y_1, Y_2) | Y_1] = E[X | Y_1]
$$

More generally,
$$
E[X|\mathcal G] = E[E(X | \mathcal H) | \mathcal G] \quad \mathcal G \subseteq \mathcal H
$$

### Properties

Assume that $\{Z_n : n \geq 1 \}$ is a martingale.
Show that for $k \in [1, n)$
$$
E[Z_n | Z_1, \ldots, Z_k] = Z_k
$$

If $n = k + 1$, then by matingale properties $E[Z_{k + 1} | Z_1, \ldots, Z_k] = Z_k$.



## References

## Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
session_info()
```
