---
title: Even number on 10 distinct rolls
draft: true
author: Stefan Eng
date: '2020-05-16'
slug: even-number-on-10-distinct-rolls
categories:
  - math
tags:
  - probability
  - exercise
output:
  blogdown::html_page:
    toc: no
    fig_width: 5
    fig_height: 5
csl: ../../static/bibtex/acm-sig-proceedings.csl
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## Load frequently used packages for blog posts
library(sessioninfo) # for session_info()
library(ggplot2)
library(here)

## For R images to display well in the RSS feed (disable for local preview)
# knitr::opts_knit$set(base.url = 'stefanengineering.com/post/')

# Don't show code by default
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, comment = NA, message = FALSE)
knitr::opts_knit$set(global.par = TRUE)

# CB Palette like ESL
cbPalette <- c(
  "#999999", "#E69F00", "#56B4E9", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# My favorite theme
theme_set(ggplot2::theme_minimal(base_size = 15, base_family = "Arial"))
```

Assume that a dice is roll until 10 even numbers have been rolled.
Denote $X_i$ as the number of rolls that land on side $i$.

### Evens 

Since we know that we have 10 even rolls then each $X_2, X_4, X_6$ are binomial with $n = 10$ and $p = 1/3$.
Thus, 
$$
E[X_2] = E[X_4] = E[X_6] = \frac{10}{3}
$$

### Odds

Let $N$ be the total number of dice rolls until 10 even numbers are rolled.
This is a negative binomial distribution with parameters $r = 10$ and $p = 1/2$.
$$
P(N = n) = {n - 1 \choose 10} (1/2)^{n} \quad \text{for }n = 10, 11, \ldots
$$
with $E[N] = 20$.

Given a fixed $N = n$, then $X_1 | N = n$ is binomial with size $n - 10$ and probability $p = 1/3$.
$$
E[X_1 | N = n] = \frac{n - 10}{3}
$$
Then we can compute $E[X_1]$ as follows
$$
\begin{aligned}
  E[X_1] &= E[E[X_1 | N = n]]\\
  &= \sum_{n = 10}^\infty E[X_1 | N = n] P(N = n)\\
  &= \sum_{n = 10}^\infty \frac{n - 10}{3} P(N = n)\\
  &= \frac{1}{3} \sum_{n = 10}^\infty n P(N = n) -  \frac{10}{3} \underbrace{\sum_{n = 10}^\infty P(N = n)}_{1}\\
  &= \frac{1}{3} E[N] - \frac{10}{3}\\
  &= \frac{20 - 10}{3}\\
  &= \frac{10}{3}
\end{aligned}
$$

```{r}
library(MASS)

ross_14 <- function(N = 10) {
  # Roll the dice
  res <- NULL
  repeat {
    res <- c(res, sample(1:6, size = 50, replace = TRUE))
    evenCount <- cumsum(res %% 2 == 0)
    ind <- which(evenCount == N)
    
    if(length(ind) > 0) break;
  }
  
  res[1:ind[1]]
}

res <- replicate(1000, sum(ross_14() == 1))

ff <- fitdistr(res, "Negative Binomial")
ffp <- fitdistr(res, "Poisson")
x <- seq(0, 20, by = 1)

hist(res, probability = TRUE, ylim = c(0, .25))
points(x + .5, dnbinom(x, size = ff$estimate["size"], mu = ff$estimate["mu"]))
points(x + .5, dpois(x, ffp$estimate), col = "red")

```

### References

### Reproducibility

```{r reproducibility, echo = FALSE}
## Reproducibility info
options(width = 120)
session_info()
```
